---
title: "Webpage presentation with Plotly"
author: "Natalia Ravinskaya"
date: "July 28, 2020"
output: ioslides_presentation
---

***

This is a web page presentation created in R Markdown. 

According to the terms of the assignment, this presentation contains the date of its creation (on the titel page).  

And it also contains two plots created with Plotly. These plots show consequences of severe weather events in the United States from 1950 to November 2011.  

For more details about the dataset and their processing see my course project <https://rpubs.com/NRavinskaya/554675> for the course "Reproducible Research".

***

Ten types of events that were most harmful with respect to population health.

```{r plot1, echo=FALSE, message=FALSE}
## 1. Across the United States, which types of events are most harmful with respect to population health?
## 2. Across the United States, which types of events have the greatest economic consequences?

## Data Processing
### Loading data into R

if(!file.exists("./repdata_data_StormData.csv.bz2")){
    fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
    download.file(fileUrl, destfile="repdata_data_StormData.csv.bz2")
}
rawdata <- read.csv("repdata_data_StormData.csv.bz2")

### Parameters of interest

library(dplyr)
data <- select(rawdata, "EVTYPE", "FATALITIES":"CROPDMGEXP")

### Non-zero observations

dataNoZero <- subset(data, FATALITIES>0 | INJURIES>0 | PROPDMG>0 | CROPDMG>0)

### Converting exponent column for property damage

dataNoZero$PROPDMGEXP <- toupper(as.character(dataNoZero$PROPDMGEXP))
uEXP <- unique(dataNoZero$PROPDMGEXP)

x <- c("","-","+","0","2","3","4","5","6","7","H","K","M","B")
y <- c(10^0,10^0,10^0,10^0,10^2,10^3,10^4,10^5,10^6,10^7,10^2,10^3,10^6,10^9)
dataE <- data.frame(x,y)
for (i in 1:nrow(dataE)) {
dataNoZero <- mutate(dataNoZero,
                     PROPDMGEXP=replace(PROPDMGEXP, PROPDMGEXP==dataE[i,1], dataE[i,2]))
}

### Converting exponent column for crop damage

dataNoZero$CROPDMGEXP <- toupper(as.character(dataNoZero$CROPDMGEXP))
uEXP <- unique(dataNoZero$CROPDMGEXP)

x <- c("","?","0","K","M","B")
y <- c(10^0,10^0,10^0,10^3,10^6,10^9)
dataE <- data.frame(x,y)
for (i in 1:nrow(dataE)) {
    dataNoZero <- mutate(dataNoZero,
                         CROPDMGEXP=replace(CROPDMGEXP, CROPDMGEXP==dataE[i,1], dataE[i,2]))
}

### Combining damage amounts

dataNoZero$PROPDMGEXP <- as.numeric(dataNoZero$PROPDMGEXP)
dataNoZero$CROPDMGEXP <- as.numeric(dataNoZero$CROPDMGEXP)

dataNoZero <- dataNoZero %>%
    mutate(PROPDMG=PROPDMG*PROPDMGEXP,
           CROPDMG=CROPDMG*CROPDMGEXP) %>%
    select(-c(PROPDMGEXP,CROPDMGEXP))

### Cleaning event types
#### Converting to lower-case and creating a list of event types

dataNoZero$EVTYPE <- tolower(as.character(dataNoZero$EVTYPE))
u <- unique(dataNoZero$EVTYPE)

#### Example of data selection

AstronomicalLowTide <- grep("low tid", u, value = T)
dataNoZero$EVTYPE <- gsub(AstronomicalLowTide, "Astronomical Low Tide", dataNoZero$EVTYPE)
data <- filter(dataNoZero, EVTYPE == "Astronomical Low Tide")
dataNoZero <- anti_join(dataNoZero,data)
u <- unique(dataNoZero$EVTYPE)

#### Creating matrix for replacement

dataM <- matrix(c("avalanc", "Avalanche",
                  "bliz", "Blizzard",
                  "surg|tid", "Storm Surge/Tide",
                  "wind chil", "Cold/Wind Chill",
                  "coastal flood", "Coastal Flood",
                  "landslid", "Debris Flow",
                  "freezing fog", "Freezing Fog",
                  "fog", "Dense Fog",
                  "dense smoke", "Dense Smoke",
                  "drought", "Drought",
                  "devil", "Dust Devil",
                  "dust", "Dust Storm",
                  "excessive heat", "Excessive Heat",
                  "extrem. cold|wind chil|windchil", "Extreme Cold/Wind Chill",
                  "cold", "Cold/Wind Chill",
                  "flash", "Flash Flood",
                  "lakeshore flood", "Lakeshore Flood",
                  "flood", "Flood",
                  "frost|freeze", "Frost/Freeze",
                  "funnel cloud", "Funnel Cloud",
                  "lightning", "Lightning",
                  "marine hail", "Marine Hail",
                  "marine thunderstorm|marine tstm", "Marine Thunderstorm Wind",
                  "torn", "Tornado",
                  "hail", "Hail",
                  "tstm|tund|thu", "Thunderstorm Wind",
                  "heat", "Heat",
                  "heavy rain|hvy|rainfall|shower", "Heavy Rain",
                  "effect|lake snow", "Lake-Effect Snow",
                  "heavy snow", "Heavy Snow",
                  "rip current", "Rip Current",
                  "surf", "High Surf",
                  "marine high wind", "Marine High Wind",
                  "marine strong wind", "Marine Strong Wind",
                  "high wind", "High Wind",
                  "hurricane|typhoon", "Hurricane (Typhoon)",
                  "seiche", "Seiche",
                  "sleet", "Sleet",
                  "ice", "Ice Storm",
                  "strong wind", "Strong Wind",
                  "waterspout", "Waterspout",
                  "depres", "Tropical Depression",
                  "tropical storm", "Tropical Storm",
                  "tsunam", "Tsunami",
                  "volcanic ash", "Volcanic Ash",
                  "fire", "Wildfire",
                  "winter storm", "Winter Storm",
                  "winter weather|freezing", "Winter Weather"),
                ncol = 2, byrow = T)

#### Processing and Transfer

for (i in 1:nrow(dataM)) {
    x <- grep(dataM[i,1], u, value = T)
    for (j in 1:length(x)) {
        dataNoZero <- mutate(dataNoZero, EVTYPE=replace(EVTYPE, EVTYPE==x[j], dataM[i,2]))
    }
    a <- filter(dataNoZero, EVTYPE == dataM[i,2])
    data <- bind_rows(data,a)
    dataNoZero <- anti_join(dataNoZero,data)
    u <- unique(dataNoZero$EVTYPE)
}

## Results
## Ten types of events that were most harmful with respect to population health.

ph <- data %>%
    group_by(EVTYPE) %>%
    summarise(dead=sum(FATALITIES),
              injured=sum(INJURIES),
              total=dead+injured) %>%
    arrange(desc(total))
ph10 <- head(ph, 10)

library(tidyr)
ph10_long <-gather(ph10, people, number, -EVTYPE)


# load packages
library(plotly)
library(ggplot2)

g1 <- ggplot(ph10_long, aes(reorder(EVTYPE, -number), number)) +
    geom_bar(stat="identity", aes(fill=people), position="dodge") +
    labs(x="Event Type", y="Frequency Count") +
    theme(axis.text.x = element_text(angle=45, hjust=1))

(gg1 <- ggplotly(g1))
```

***

Ten types of events that had the greatest economic consequences.

```{r plot2, echo=FALSE, message=FALSE}
## Ten types of events that had the greatest economic consequences.

ec <- data %>%
    group_by(EVTYPE) %>%
    summarise(property=sum(PROPDMG),
              crop=sum(CROPDMG),
              total=property+crop) %>%
    arrange(desc(total))
ec10 <- head(ec, 10)

ec10_long <-gather(ec10, damage, amount, -EVTYPE)

g2 <- ggplot(ec10_long, aes(reorder(EVTYPE, -amount), amount)) +
    geom_bar(stat="identity", aes(fill=damage), position="dodge") +
    labs(x="Event Type", y="Frequency Count") +
    theme(axis.text.x = element_text(angle=45, hjust=1))

(gg2 <- ggplotly(g2))
```


